---
title: "Math 660"
output:
  pdf_document: default
  html_document: default
date: "December 2022"
theme: cerulean
---

<!-- For more info on RMarkdown see http://rmarkdown.rstudio.com/ -->

## Name: Emiliana Geronimo

### 1. In my simulation, I created the empirical and standard normal CDF. Firstly, my code used lambda of 10, a sample size of 2200, and a random number generator(rpois) to create data. Then with that data, I coded the equation to find Xbar. This was done 1000 times in my simulation to obtain 1000 different Xbars from 1000 different datasets. Then, I used the ecdf function to find the empirical CDF of the Xbars and plotted it. To confirm the shape of the CDF, I made a sequence of 100 points from -3 to 3 and used pnorm (gives vector of probabilites) to find the normal CDF. Overall, the points depicts the cumulative distribution function of the Xbars in my simulation. The CDF only approaches 0 to 1 because it is the probability that a random variable of X (pdf) with a given probability distribution will be discovered at a value less than or equal to that X which is shown in the left side of CLT. 
```{r, fig.align='center'}
n <- 2200
lambda <- 10 
std.X.bar <-  NULL 
for (r in 1:1000) {
  X <- rpois(n,lambda)
  std.X.bar[r] <- sqrt(n)*(mean(X)-lambda)/sqrt(lambda)
}

p <- ecdf(std.X.bar)
plot(p, main= "Empirical and Standard Normal CDF", xlab= "Xbar")
x <- seq(-3,3,length=100)
y <- pnorm(x) 
lines(x,y,col='red')

```

### 2. The Newton-Pepys problem debated the probability of throwing a number of sixes from a certain number of dice. The answer to this question could be found using binomial distributions. However, to answer this question, I created a simulation that passed in the number of dice and the number of sixes to find the probabilites of Scenarion A, B, and C. I first created a sample (sample size of whatever dice is inputed) of numbers 1-6 with replacement. Then, I examined in each sample for n amount times, how many times the six showed up, then found the proportions for each scenario. Pepys claimed that Scenario C (of 18 dice and 3 sixes) had a greater chance of success, however he is wrong. The greatest chance of success amongst these scenarios is Scenario A (of 6 dice and 1 six). According to the solution of the Newton-Pepys problem, my simulation was supports the probabilty of Scenario A > Scenario B > Scenario C and portrayed the binomial distribution properly.

```{r}
diceroll <- function(dice,a) {
  n <- 2000
  b <- 0
  for (i in 1:n) {
    dsample <- sample(1:6, size=dice, replace=TRUE)
    count = 0 
    for (d in 1:dice) {
      if (dsample[d]==6) {
        count <- count + 1
      }
    }
    if (count >= a) {
      b <- b + 1
    }
  }
  return(b/n)
}
```
```{r}
diceroll(6,1)
diceroll(12,2)
diceroll(18,3)
```

### 3. This function creates the inverse CDF and plots it with the exponential pdf. The inverse CDF is x = –log(1–u). My code creates random values from the exponential distribution by generating random uniform values from U(0,1) and applying the inverse CDF of the exponential.
```{r, echo=FALSE, fig.align='center'}
function3 <- function(n) {
  XX <- NULL
  for (i in 1:1000) {
    UU <- runif(n)
    XX[i] <- -log(1-UU[i])
  }
  return(XX)
}

bb <- function3(1000)
hist(bb, freq=FALSE, main = "Histogram of Inverse CDF")
f <-seq(0,10,length=100)
k <-dexp(f)
lines(f,k,col="red")
```

